Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics Experiments

### 2013-11-19 papers/2013/NSS/NSS.tex
Fixed some typos

\subsection{Code implementation}
The starting point of of our studies is the SVT track fitting
simulation code, written in the \textsc{c} language. SVT track fitting
is divided into three main functions: first, the unpacking of input
data and filling of all the necessary data structures; second, the
computation of all possible combinations of hits in each road and
third, the linearized track fit of each combination of hits. Three
main loops are present - on events, roads and hit combinations.

@@

To measure the data transfer latency we use a computing cluster composed of 
12 identical nodes.  Each node contains a Intel Xeon E6520 2.4 GHz 
CPU and two Tesla M2075 GPU cards. The nodes are connected by InfiniBand 
communication links using Connect-X2 Mellanox or APEnet+ adapters. 
APEnet+ is an FPGA-based PCIe board supporting peer-to-peer communication with 
Tesla and Kepler cards~\cite{apenet2010}.
Two nodes of this cluster are used to measure data transfer latency,
one acting as a transmitter and the
other as a receiver.  Data are transferred from the transmitter to the 
receiver, processed on the GPU and sent back to the receiver (see
Fig.~\ref{fig:data_flow}).  
The latency for a complete loop is measured on the transmitter using 
standard \textsc{c} libraries.
\begin{figure}[tbp]
\centering
\includegraphics[width=3.5in]{figures/SetUp-general}
\caption{Data flow. Data is sent from the transmitter PC to the
  receiver PC, where it is processed by the GPU before being returned
  to the transmitter PC. The transmitter plays the role of the
  detector as the source of the data and as an upstream trigger
  processor as the data's ultimate sink. The receiver PC plays the
  role of a component in the trigger system. }
\label{fig:data_flow}
\end{figure}
In this setup, the transmitter can represent the detector, as
the source of the data, or an upstream trigger processor, as
the ultimate sink of the data, while the the receiver is the
trigger system: the time to transfer data to the receiver is thus a
rough estimate of the latency to transfer the data from the detector
front-end to the trigger system.

@@

\subsection{Data processing}
Each data sample is processed 100 times by the track fitting algorithm. 
The average latency as a function of the number of fits is presented in 
Fig.~\ref{fig:algo_only_timing} for the serial,
embarrassingly parallel and parallel algorithms. We see that the
embarrassingly parallel algorithm gives a modest increase with respect
to the serial (CPU) algorithm. Switching to a fully parallel algorithm
affords a much more significant speed improvement. 
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TimeComp_MIC}
  \caption{Algorithm-only comparison for timing as a function of the
    number of track fits. We compare timing on CPUs (serial), Intel
    MIC (embarrassingly parallel), and GPUs (fully parallel), in
    blue, greed, and red, respectively. The GPUs exhibit the best
    performance due to the full parallelization. }
  \label{fig:algo_only_timing}
\end{figure}
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TimeCompZoom_MIC.png} 
  \caption{Algorithm-only comparison for timing as a function of the
    number of track fits: zoom in the low number of fits region. At
    low number of fits, the CPU performs better, due to start-up
    costs associated with data transfers to the accelerator card.}
  \label{fig:algo_only_timing_zoom}
\end{figure}
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC}
  \caption{Speed-up with respect to a standard CPU (Intel Xeon
    E5630). The speed-ups plateau after about two million fits.}
  \label{fig:algo_only_speedup}
\end{figure}
The accelerator card's performance drop with decreasing number of fits, as can be seen 
in Fig.~\ref{fig:algo_only_timing_zoom}, due to overhead.
Figure~\ref{fig:algo_only_speedup} shows the speed-up with respect to the serial 
algorithm run on a standard CPU (Intel Xeon E5630): the maximum gain is 
obtained processing at least 500 events. This means that to fully exploit 
parallel architectures millions of fits have to be performed in parallel.

>>>>

\subsection{Code implementation}
The starting point of our studies is the SVT track fitting
simulation code, written in the \textsc{c} language. SVT track fitting
is divided into three main functions: first, the unpacking of input
data and filling of all the necessary data structures; second, the
computation of all possible combinations of hits in each road and
third, the linearized track fit of each combination of hits. Three
main loops are present - on events, roads and hit combinations.

@@

To measure the data transfer latency we use a computing cluster composed of 
12 identical nodes.  Each node contains a Intel Xeon E6520 2.4 GHz 
CPU and two Tesla M2075 GPU cards. The nodes are connected by InfiniBand 
communication links using Connect-X2 Mellanox or APEnet+ adapters. 
APEnet+ is an FPGA-based PCIe board supporting peer-to-peer communication with 
Tesla and Kepler cards~\cite{apenet2010}.
Two nodes of this cluster are used to measure data transfer latency,
one acting as a transmitter and the
other as a receiver.  Data are transferred from the transmitter to the 
receiver, processed on the GPU and sent back to the receiver (see
Fig.~\ref{fig:data_flow}).  
The latency for a complete loop is measured on the transmitter using 
standard \textsc{c} libraries.
\begin{figure}[tbp]
\centering
\includegraphics[width=3.5in]{figures/SetUp-general}
\caption{Data flow. Data is sent from the transmitter PC to the
  receiver PC, where it is processed by the GPU before being returned
  to the transmitter PC. The transmitter plays the role of the
  detector as the source of the data and as an upstream trigger
  processor as the data's ultimate sink. The receiver PC plays the
  role of a component in the trigger system. }
\label{fig:data_flow}
\end{figure}
In this setup, the transmitter can represent the detector, as
the source of the data, or an upstream trigger processor, as
the ultimate sink of the data, while the receiver is the
trigger system: the time to transfer data to the receiver is thus a
rough estimate of the latency to transfer the data from the detector
front-end to the trigger system.

@@

\subsection{Data processing}
Each data sample is processed 100 times by the track fitting algorithm. 
The average latency as a function of the number of fits is presented in 
Fig.~\ref{fig:algo_only_timing} for the serial,
embarrassingly parallel and parallel algorithms. We see that the
embarrassingly parallel algorithm gives a modest increase with respect
to the serial (CPU) algorithm. Switching to a fully parallel algorithm
affords a much more significant speed improvement. 
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TimeComp_MIC}
  \caption{Algorithm-only comparison for timing as a function of the
    number of track fits. We compare timing on CPUs (serial), Intel
    MIC (embarrassingly parallel), and GPUs (fully parallel), in
    blue, green, and red, respectively. The GPUs exhibit the best
    performance due to the full parallelization. }
  \label{fig:algo_only_timing}
\end{figure}
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TimeCompZoom_MIC.png} 
  \caption{Algorithm-only comparison for timing as a function of the
    number of track fits: zoom in the low number of fits region. At
    low number of fits, the CPU performs better, due to start-up
    costs associated with data transfers to the accelerator card.}
  \label{fig:algo_only_timing_zoom}
\end{figure}
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/Speedup_MIC}
  \caption{Speed-up with respect to a standard CPU (Intel Xeon
    E5630). The speed-ups plateau after about two million fits.}
  \label{fig:algo_only_speedup}
\end{figure}
The accelerator card's performance drop with decreasing number of fits, as can be seen 
in Fig.~\ref{fig:algo_only_timing_zoom}, due to overhead.
Figure~\ref{fig:algo_only_speedup} shows the speed-up with respect to the serial 
algorithm run on a standard CPU (Intel Xeon E5630): the maximum gain is 
obtained processing at least 500 events. This means that to fully exploit 
parallel architectures millions of fits have to be performed in parallel.
